### 2025-04-10 chapters/5-event_shapes.tex
Final changes to main files for first draft

In the context of pair-wise correlations, \glspl{ewoc} can be defined as a sum of pair-wise contributions:
\begin{align}
    \label{eq:intro_ewoc_def}
    \frac{\dd \Sigma_\mathcal{O}}{\dd \chi}
    =
    \frac{1}{\sigma}
    \int \dd\sigma \,
    \sum_{
        i,\, j
    } \,
    z_i \, z_j \,
    \delta\left(\chi \, - \, \mathcal{O}_{ij}\right)
    ,
\end{align}
which yields a normalized probability distribution in \(\chi\).
%
\((\dd)\sigma\) indicates the (differential) cross section for jet production in the process under consideration, and the energy-weighting factors \(z_{i,j}\) ensure that higher-energy objects have a greater impact on the result of \Eq{intro_ewoc_def}.%
\footnote{
    For proton-proton collisions at the LHC, transverse momentum fractions \(z_i = p_{T,i}/p_{T,\text{jet}}\) are a natural choice, while energy fractions \(z_{i} = E_{i}/E_\text{jet}\) are more natural in the context of electron-positron collisions.
}
%
\(\mathcal{O}_{ij}\) is a user-defined  observable on pairs of subjets, such as their angular separation or invariant mass.
%
An \gls{ewoc} involving pair-wise contributions is also called a \vocab{two-point} \gls{ewoc}
%
A more general definition of \glspl{ewoc} involving higher-point correlations and more will be given in \Sec{ewoc}.

@@

Among jet substructure observables, the Energy-Energy Correlator (EEC) \cite{Basham:1978bw,Basham:1978zq,Basham:1979gh} is particularly noteworthy:
%
it is conceptually simple, insensitive to low-energy radiation, and naturally separates physics at different scales.
%
Physically, the EEC measures the mean correlation in the energy passing through any two detectors separated by an angle \(\Delta\), averaged over many particle collisions:
%
\(
    \text{EEC}(\Delta)
    \propto
    \left\langle
        E_\text{det.\.1} \,\,\, E_\text{det.\.2}
    \right\rangle
\).%
\footnote{
    More concretely, this means that the contribution of each particle in an event to the EEC is weighted by the particle's energy,
    %
    see \Eq{eec_defn}.
    %
    This energy weighting mitigates the effects of the low-energy physics of QCD that complicates the theoretical description of other jet substructure observables.
}
%
From the birth of the EEC to its use in recent collider studies, the simplicity of the EEC has enabled highly accurate perturbative computations \cite{Clay:1995sd,Glover:1994vz,Kramer:1996qr,DelDuca:2016csb,Gituliar:2017umx,Dixon:2018tpg,Dixon:2018qgp,Henn:2019gkr,Luo:2019nig,Gao:2020vyx,Neill:2022lqx,Lee:2023npz,Kramer:1995qh,deFlorian:2004mp,Banfi:2002vw,Tulipant:2017ybb,Moult:2018jzp,Korchemsky:2019nzm,Dixon:2019uzg,Gao:2019ojf,Luo:2019hmp,Luo:2019bmw,Moult:2019vou,Li:2020bub,Ebert:2020sfi,Li:2021txc,Duhr:2022yyp,Chen:2023zlx,Gao:2023ivm} and has facilitated measurements of the strong coupling constant \(\alpha_s\) \cite{Martin:1986uq,DELPHI:1990sof,SLD:1994yoe,ATLAS:2015yaa,ATLAS:2017qir,dEnterria:2018cye,Kardos:2018kqj,Ali:2020ksn,dEnterria:2022hzv,ATLAS:2023tgo} including the most precise jet substructure determinations of \(\alpha_s\) to date \cite{CMS:2024mlf}.
%
The EEC has been investigated as a probe of jets traveling through the quark-gluon plasma (QGP) produced in heavy-ion collisions \cite{Lokhtin:2004tx,Lokhtin:2006dp,Andres:2022ovj,Barata:2023zqg,Andres:2023xwr,Yang:2023dwc,Barata:2023vnl,Barata:2023bhh,Barata:2024nqo} and of correlations in nuclear physics \cite{Karapetyan:2019fst,Liu:2022wop,Liu:2023aqb,Kang:2023gvg,Cao:2023oef}.
%
The study of the EEC has also shed light on the structure of non-perturbative QCD \cite{Nason:1995np,Korchemsky:1997sy,Korchemsky:1999kt,Dokshitzer:1999sh,Chen:2020vvp,Jaarsma:2023ell,Schindler:2023cww,Lee:2024esz}
and of quantum field theory in general \cite{Richards:1983sr,Sveshnikov:1995vi,Hofman:2008ar,Hatta:2012kn,Belitsky:2013bja,Belitsky:2013ofa,Belitsky:2013xxa,Belitsky:2014zha,Korchemsky:2015ssa,Goncalves:2014ffa,Farnsworth:2015hum,Hofman:2016awc,Kravchuk:2018htv,Kologlu:2019mfz,Chang:2020qpj,Korchemsky:2021htm,Caron-Huot:2022eqs,Chen:2023wah,Chicherin:2023gxt,Chicherin:2024ifn}.
%
It also been put forward as a way to search for new physics in hadronic decays of electroweak bosons~\cite{Ricci:2022htc}.
%
Finally, while most jet substructure observables are dominated by contributions from two-particle correlations, the EEC can be naturally extended to multi-point energy correlators which can characterize correlations between an arbitrary number of particles.
%
Recent theoretical work suggests that multi-point generalizations of the EEC may be used to extract the mass of the top quark at the LHC \cite{Procura:2022fid,Holguin:2022epo,Holguin:2023bjf,Pathak:2023tmy,Xiao:2024rol,Holguin:2024tkz}.

@@

\remark{eec-energyflow}{
    The EEC of \Eq{eec-real-angle} in a theory with massless particles (or in the high-energy limit \(m^2 / Q^2 \ll 1\) for all particle masses \(m\)) can also be expressed in terms of the energy flow operators of \Def{energy-flow} as
    \begin{align}
        \label{eq:eec-energy-flow}
        \frac{\dd \Sigma_\text{eec}}{\dd\chi}
        =
        \int
        \,
        \dd\Omega_1 \dd\Omega_2
        \,
        \le\langle
            \frac{1}{Q^2}
            \mathcal{E}(\hat{n}_1)
            \mathcal{E}(\hat{n}_2)
        \ri\rangle
        \,\,
        \delta\le(
            \chi - \frac{1 - \hat{n}_1\cdot\hat{n}_2}{2}
        \ri)
        \,,
    \end{align}
    where the expectation value indicates a quantum mechanical average, potentially over an ensemble of initial states.
    %
    When measuring correlations of distance \(\Delta R\) in the rapidity-azimuth plane -- i.e. in proton-proton collisions -- the delta function of \Eq{eec-energy-flow} must be replaced by
    \begin{align}
        \delta\le(R_{ij} - \Delta R\le(\hat{n}_1, \hat{n}_2\ri)\ri)
        \,.
    \end{align}
    In practice, one will also add cuts on the radiation considered in the EEC or consider the EEC only within a specific jet, which requires the use of additional constraints (for example, the use of Heaviside Theta functions in the integrand, dependent on the energy flow itself, which enforce that \(\hat{n}_1\) and \(\hat{n}_2\) lie within the same jet).
    %
    While these additional constraints may be complicated functionals of the energy flow operator, they may generally be disregarded in the collinear limit that we consider.
}

@@

\begin{proposition}{The (differential) NLO EEC Jet Function}{nlo-eec-jet-function}
The differential EEC for any process containing a massless intermediate quark (or anti-quark) contains a universal piece of the form
\begin{align}
    \label{eq:universal_differential_eec}
    \frac{\dd \Sigma_\text{EEC}}{\dd \chi}
    (\chi \to 0)
    \supset
    \le[
    \frac{3\ascf}{\chi}
    \ri]_+
    \,,
\end{align}
up to a numerical, process dependent coefficient encoding the average number density and energy fraction of quarks (or anti-quarks).

\Eq{universal_differential_eec} is the EEC associated with a single splitting within a quark jet, normalized to the energy of the quark.
%
Therefore, it is the \(\mathcal{O}(\alpha_s)\) (NLO) jet function for the \textit{differential} EEC \cite{Dixon:2019uzg}.
\end{proposition}

@@

We conclude that the contribution of a single jet (or initiating parton) to the EEC in the collinear limit can be separated into two distinct, calculable pieces (see also \Fig{shower:cartoon:collinear2}):
\begin{enumerate}
    \item[a)]
    \textbf{The penultimate parton:}
    %
    parton-to-parton fragmentation produces one final intermediate/virtual parton \(f\) according to the partonic fragmentation functions \Eq{fragmentation_function}
    %
    (see also \Fig{shower:cartoon:fragmentation}).
    %
    We derived the relevant expressions for parton-to-parton fragmentation in \Sec{p2p-fragmentation}, where we found that the Mellin moments of parton-to-parton fragmentation functions were given by \Eq{ll_fragmentation_function}.
    %
    We repeat them here with \(\theta_f = 2\sqrt{\chi}\) (from \(\chi = (1-\cos\theta)/2, \theta\ll1\)) for convenience:
    \begin{align*}
        \hat{F}_{f \leftarrow i} (m;\,
        \theta_f = 2 \sqrt{\chi}\leftarrow \Rjet)
        =
        \le(
            \alpha_s\le(2\,Q\sqrt{\chi}\ri)
            \, / \,
            \alpha_s\le(Q\,\Rjet\ri)
        \ri)
        ^
        {\hat{p}(m) \, / \, 2\pi\beta_0}
        _{f\leftarrow i}
        ;
    \end{align*}

@@

Futhermore, since \Eq{jet_fn_ll_0} gives the (normalized) contribution of a narrow jet of partonic flavor \(i\) to the EEC in the collinear limit, it is precisely the jet function of \Eq{jet_fn_defn}.
%
Using \Eqss{ll_fragmentation_function}{splitting_fn_contribution_to_eec}{jet_fn_ll_0}, we can then write the LL jet function (EEC for a narrow jet of flavor \(i\)) as
\begin{align}
    \label{eq:eec_jet_fn_ll_1}
    J^i(\chi; \, Q)
    &=
    \frac{2 \as}{\chi}
    \,
    \sum_{a,b,\,f}
    \le(
        \int_0^1 \dd z \, z (1-z) \,
        P_{ab \leftarrow f}(z)
    \ri)
    \,
    \le(
        \frac{\alpha_s\le(2\,Q\sqrt{\chi}\ri)}
        {\alpha_s\le(Q\,\Rjet\ri)}
    \ri)
    ^
    {\frac{\hat{p}(2)}{2\pi\beta_0}}
    _{f\leftarrow i}
    .
\end{align}

@@

Unfortunately, the expression of \Eq{old_def} is computationally expensive due to the intricate phase space constraint \(R_L = \max_{k,\ell}\{R_{i_k,i_\ell}\}\).
%
For example, the time required to compute the (integer) \gls{penc} scales as $M^N/N!$ for a jet with $M$ particles; when analytically continued to non-integer values of $N$, the \gls{penc} suffers  a computational scaling of $2^{2M}$.%
%
\footnote{
    \textsc{FastEEC}~\cite{Budhraja:2024xiq} achieves a substantial speed up by replacing particles with subjets whose radius is chosen dynamically, depending on the desired level of angular resolution.
    %
    For non-integer $N$, \textsc{FastEEC} furthermore uses a recursive algorithm that reduces the standard $2^{2M}$ scaling down to $M\,2^M$~\cite{Budhraja:2024tev}.
}
%
This computational cost impedes several exciting \gls{penc} applications.
%
For example, \glspl{penc} have an approximate scaling behavior related to the $N$-th Mellin moment of the DGLAP splitting functions~\cite{Chen:2020vvp};
%
the analytic continuation of the \gls{penc} to non-integer $N$ therefore provides access both to the full splitting functions of QCD and, in the limit $N \to 0$, a unique opportunity to study small-$x$ physics and BFKL dynamics with jets~\cite{Chen:2020vvp, Budhraja:2024tev,Neill:2020bwv}.
%
At very large values of \(N\), \glspl{penc} also encode additional fundamental features of QCD~\cite{Chen:2020vvp,Dai:2024wff}, such as level crossings with twist-4 operators.
%
Furthermore, in the high-multiplicity environment of heavy ion collisions, even computing \glspl{penc} for \(N=3\) has been computationally challenging, limiting their potential in studying medium effects~\cite{Bossi:2024qho}.
%
Improved computational efficiency is therefore necessary to leverage the full potential of \glspl{penc} in the study of realistic data samples.

Therefore, rather than focusing on the traditional projections of ENCs, in this section we focus on new parametrizations of higher point energy correlators, introduced in \Reff{Alipour-fard:2024szj}, with a number of improved properties.
%
First, our parametrization of the \gls{penc} depends on the largest distance \(R_1\) to a ``special'' particle $s$ in a set of $N$ particles, suitably averaged over all choices for $s$;
%
this yields simpler phase space restrictions than the traditional parametrization for the \gls{penc} in terms of the largest pairwise angle~\cite{Chen:2020vvp}.
%
Second, when considering more differential information, our parametrization of resolved ENCs (\gls{renc}s) employs non-redundant polar coordinates centered around the special particle, as in \Fig{cartoon}, yielding intuitive visualizations of jet substructure.
%
This differs from the traditional approach, which uses over-complete information from the set of all pairwise distances and neglects information about the relative orientation of particles.
%
Third, our parametrization offers dramatic improvements in computational performance, which is essential for their use in experimental analyses.
%
Finally, we anticipate that these conceptual and computational improvements will yield simpler theoretical calculations.
%
Our implementation for computing the \glspl{penc} of \Eq{new_def} be found on GitHub as an update to \textsc{FastEEC}~\cite{FASTEEC}, and of our \glspl{penc} and \gls{renc}s at \texttt{ResolvedEnergyCorrelators}~\cite{github:RENC}.

@@

EWOCs therefore trade the ability to capture correlations between all particles in a jet for the ability to more easily control the effects of low-energy QCD.
%
As we will see, this tradeoff dramatically increases the richness of the correlations that can be probed by EWOCs.
As we discuss further in \Sec{ewoc-irc}, EWOCs are not collinear safe without an infra-red regulator;
%
therefore, we consider EWOCs which capture correlations between \textit{subjets}.%
\footnote{
    In principle, we could use any infra-red regulator -- see Remark~\ref{lund-eec}.
}
%
These subjet EWOCs, which we will henceforth simply call EWOCs, have some important similarities and differences from the energy correlators defined above.

In addition to the angular correlations probed by the EEC, EWOCs can be used to capture many other correlations -- such as correlations in mass, formation time, or any other observable on subjets -- which are forbidden by collinear safety in the absence of a subjet radius.
%
An EWOC for a generic observable \(\mathcal{O}(\text{subjet 1, subjet 2})\) on pairs of subjets takes a similar form to \Eq{eec_defn}.
%
Its definition was given in \Eq{intro_ewoc_def} and is repeated here for convenience:
%
\begin{definitionbox}{Pairwise EWOC}{pairwise-ewoc}
    A \glslink{ewoc}{\vocab{pairwise (or two-point) EWOC}} takes the form
    \begin{align}
        \label{eq:intro_ewoc_def}
        \frac{\dd \Sigma_\mathcal{O}}{\dd \chi}
        =
        \frac{1}{\sigma}
        \int \dd\sigma \,
        \sum_{
            \text{subjets }
            i,\, j
        } \,
        z_i \, z_j \,
        \delta\left(\chi \, - \, \mathcal{O}_{ij}\right)
        ,
    \end{align}
    where \((\dd)\sigma\) indicates the (differential) cross section for jet production in the process under consideration and the sum on \(i\) and \(j\) run over the subjets within the jet.
    %
    The energy-weighting factors \(z_{i,j}\), adopted from the EEC, ensure that higher-energy subjets contribute more to the EWOC of \Eq{intro_ewoc_def}.%
    \footnote{
        For proton-proton collisions at the LHC, transverse momentum fractions \(z_i = p_{T,i}/p_{T,\text{jet}}\) are a natural choice, while energy fractions \(z_{i} = E_{i}/E_\text{jet}\) are more natural in the context of electron-positron collisions.
    }
    %
    \(\mathcal{O}_{ij}\) is a user-defined  observable on pairs of subjets, such as their angular separation or invariant mass.
\end{definitionbox}

@@

\remark{}{
    For example, using the two-particle angular separation in \Eq{intro_ewoc_def}, \(\mathcal O_{ij} \to \theta_{ij}\), recovers the traditional EEC of $e^+e^-$ collisions in the limit of zero subjet radius (i.e.~where the sum over pairs of subjets is replaced by a sum over pairs of particles).
}

@@

The advantage of EWOCs is that they directly probe correlations in an observable of interest.
%
The choice of subjet definition (e.g.~radius and recombination scheme) determines the set of subjets that contribute to the EWOC, and therefore may also be tuned to capture different physical effects.
%
For example, the subjet radius may be chosen to capture the subjets emerging from the hadronic decays of a boosted particle, as we explore in \Sec{mass}.
%
In this section, we focus on the concrete example of the mass EWOC as a proof-of-concept, with a subjet radius chosen to probe decays of the \(W\) boson, but the use of EWOCs is by no means limited to the measurement of mass correlations.
%
Possible pairwise observables include:
\begin{itemize}
    \item
        The opening angle \(\theta_{ij}\) or the rapidity-azimuth distance \(R_{ij}\) between subjets, which probe angular correlations (used in $e^+e^-$ or $pp$ collisions, respectively).
        %
        These observables lead to the EEC, for which the use of subjets does not provide novel results;%
        \footnote{
            With a subjet radius \rsub{}, the value of the subjet EEC is nearly identical to the particle-level EEC for $\chi>\rsub$, while the collinear cutoff of the subjet radius sets the subjet EEC to zero for $0<\chi<\rsub$.
        }

    \item
        The mass \(m_{ij} = \sqrt{\le(p_i + p_j\ri)^2}\), which probes mass scales%
        \footnote{
            The four-momenta of the subjets depends on the recombination scheme used to define the subjets.
            %
            In this section, we use winner-take-all schemes which enforce that all subjets are massless.
            %
            We have found that this choice of recombination scheme does not affect the performance of the mass EWOC in constraining the $W$-boson mass.
        }, applied to the phenomenolgy of massive decaying particles in \Sec{mass};

>>>>

In the context of pair-wise correlations, \glspl{ewoc} can be defined as a sum of pair-wise contributions:
\begin{align}
    \label{eq:intro_ewoc_def}
    \frac{\dd \Sigma_\mathcal{O}}{\dd \chi}
    =
    \frac{1}{\sigma}
    \int \dd\sigma \,
    \sum_{
        i,\, j
    } \,
    z_i \, z_j \,
    \delta\left(\chi \, - \, \mathcal{O}_{ij}\right)
    ,
\end{align}
which yields a normalized probability distribution in \(\chi\).
%
\((\dd)\sigma\) indicates the (differential) cross section for jet production in the process under consideration, and the energy-weighting factors \(z_{i,j}\) ensure that higher-energy objects have a greater impact on the result of \Eq{intro_ewoc_def}.%
\footnote{
    For proton-proton collisions at the LHC, transverse momentum fractions \(z_i = p_{T,i}/p_{T,\text{jet}}\) are a natural choice, while energy fractions \(z_{i} = E_{i}/E_\text{jet}\) are more natural in the context of electron-positron collisions.
}
%
\(\mathcal{O}_{ij}\) is a user-defined  observable on pairs of subjets, such as their angular separation or invariant mass.
%
An \gls{ewoc} involving pair-wise contributions is also called a \vocab{two-point} \gls{ewoc}
%
A more general definition of \glspl{ewoc} involving higher-point correlations and more will be given in \Sec{ewocs}.

@@

Among jet substructure observables, the Energy-Energy Correlator (EEC) \cite{Basham:1978bw,Basham:1978zq,Basham:1979gh} is particularly noteworthy:
%
it is conceptually simple, insensitive to low-energy radiation, and naturally separates physics at different scales.
%
Physically, the EEC measures the mean correlation in the energy passing through any two detectors separated by an angle \(\Delta\), averaged over many particle collisions:
%
\(
    \text{EEC}(\Delta)
    \propto
    \left\langle
        E_\text{det.\.1} \,\,\, E_\text{det.\.2}
    \right\rangle
\).%
\footnote{
    More concretely, this means that the contribution of each particle in an event to the EEC is weighted by the particle's energy,
    %
    as in \Eq{intro_ewoc_def} and, shorty, \Def{eec}.
    %
    This energy weighting mitigates the effects of the low-energy physics of QCD that complicates the theoretical description of other jet substructure observables.
}
%
From the birth of the EEC to its use in recent collider studies, the simplicity of the EEC has enabled highly accurate perturbative computations \cite{Clay:1995sd,Glover:1994vz,Kramer:1996qr,DelDuca:2016csb,Gituliar:2017umx,Dixon:2018tpg,Dixon:2018qgp,Henn:2019gkr,Luo:2019nig,Gao:2020vyx,Neill:2022lqx,Lee:2023npz,Kramer:1995qh,deFlorian:2004mp,Banfi:2002vw,Tulipant:2017ybb,Moult:2018jzp,Korchemsky:2019nzm,Dixon:2019uzg,Gao:2019ojf,Luo:2019hmp,Luo:2019bmw,Moult:2019vou,Li:2020bub,Ebert:2020sfi,Li:2021txc,Duhr:2022yyp,Chen:2023zlx,Gao:2023ivm} and has facilitated measurements of the strong coupling constant \(\alpha_s\) \cite{Martin:1986uq,DELPHI:1990sof,SLD:1994yoe,ATLAS:2015yaa,ATLAS:2017qir,dEnterria:2018cye,Kardos:2018kqj,Ali:2020ksn,dEnterria:2022hzv,ATLAS:2023tgo} including the most precise jet substructure determinations of \(\alpha_s\) to date \cite{CMS:2024mlf}.
%
The EEC has been investigated as a probe of jets traveling through the quark-gluon plasma (QGP) produced in heavy-ion collisions \cite{Lokhtin:2004tx,Lokhtin:2006dp,Andres:2022ovj,Barata:2023zqg,Andres:2023xwr,Yang:2023dwc,Barata:2023vnl,Barata:2023bhh,Barata:2024nqo} and of correlations in nuclear physics \cite{Karapetyan:2019fst,Liu:2022wop,Liu:2023aqb,Kang:2023gvg,Cao:2023oef}.
%
The study of the EEC has also shed light on the structure of non-perturbative QCD \cite{Nason:1995np,Korchemsky:1997sy,Korchemsky:1999kt,Dokshitzer:1999sh,Chen:2020vvp,Jaarsma:2023ell,Schindler:2023cww,Lee:2024esz}
and of quantum field theory in general \cite{Richards:1983sr,Sveshnikov:1995vi,Hofman:2008ar,Hatta:2012kn,Belitsky:2013bja,Belitsky:2013ofa,Belitsky:2013xxa,Belitsky:2014zha,Korchemsky:2015ssa,Goncalves:2014ffa,Farnsworth:2015hum,Hofman:2016awc,Kravchuk:2018htv,Kologlu:2019mfz,Chang:2020qpj,Korchemsky:2021htm,Caron-Huot:2022eqs,Chen:2023wah,Chicherin:2023gxt,Chicherin:2024ifn}.
%
It also been put forward as a way to search for new physics in hadronic decays of electroweak bosons~\cite{Ricci:2022htc}.
%
Finally, while most jet substructure observables are dominated by contributions from two-particle correlations, the EEC can be naturally extended to multi-point energy correlators which can characterize correlations between an arbitrary number of particles.
%
Recent theoretical work suggests that multi-point generalizations of the EEC may be used to extract the mass of the top quark at the LHC \cite{Procura:2022fid,Holguin:2022epo,Holguin:2023bjf,Pathak:2023tmy,Xiao:2024rol,Holguin:2024tkz}.

@@

\remark{eec-energyflow}{
    The EEC of \Def{eec} in a theory with massless particles (or in the high-energy limit \(m^2 / Q^2 \ll 1\) for all particle masses \(m\)) can also be expressed in terms of the energy flow operators of \Def{energy-flow} as
    \begin{align}
        \label{eq:eec-energy-flow}
        \frac{\dd \Sigma_\text{eec}}{\dd\chi}
        =
        \int
        \,
        \dd\Omega_1 \dd\Omega_2
        \,
        \le\langle
            \frac{1}{Q^2}
            \mathcal{E}(\hat{n}_1)
            \mathcal{E}(\hat{n}_2)
        \ri\rangle
        \,\,
        \delta\le(
            \chi - \frac{1 - \hat{n}_1\cdot\hat{n}_2}{2}
        \ri)
        \,,
    \end{align}
    where the expectation value indicates a quantum mechanical average, potentially over an ensemble of initial states.
    %
    When measuring correlations of distance \(\Delta R\) in the rapidity-azimuth plane -- i.e. in proton-proton collisions -- the delta function of \Eq{eec-energy-flow} must be replaced by
    \begin{align}
        \delta\le(R_{ij} - \Delta R\le(\hat{n}_1, \hat{n}_2\ri)\ri)
        \,.
    \end{align}
    In practice, one will also add cuts on the radiation considered in the EEC or consider the EEC only within a specific jet, which requires the use of additional constraints (for example, the use of Heaviside Theta functions in the integrand, dependent on the energy flow itself, which enforce that \(\hat{n}_1\) and \(\hat{n}_2\) lie within the same jet).
    %
    While these additional constraints may be complicated functionals of the energy flow operator, they may generally be disregarded in the collinear limit that we consider.
}

@@

\begin{proposition}{The (differential) NLO EEC Jet Function}{nlo-eec-jet-function}
The differential EEC for any process containing a massless intermediate quark (or anti-quark) contains a universal piece of the form
\begin{align}
    \label{eq:universal_differential_eec_prop}
    \frac{\dd \Sigma_\text{EEC}}{\dd \chi}
    (\chi \to 0)
    \supset
    \le[
    \frac{3\ascf}{\chi}
    \ri]_+
    \,,
\end{align}
up to a numerical, process dependent coefficient encoding the average number density and energy fraction of quarks (or anti-quarks).

\Eq{universal_differential_eec_prop} is the EEC associated with a single splitting within a quark jet, normalized to the energy of the quark.
%
Therefore, it is the \(\mathcal{O}(\alpha_s)\) (NLO) jet function for the \textit{differential} EEC \cite{Dixon:2019uzg}.
\end{proposition}

@@

We conclude that the contribution of a single jet (or initiating parton) to the EEC in the collinear limit can be separated into two distinct, calculable pieces (see also \Fig{shower:cartoon:collinear2}):
\begin{enumerate}
    \item[a)]
    \textbf{The penultimate parton:}
    %
    parton-to-parton fragmentation produces one final intermediate/virtual parton \(f\) according to the partonic fragmentation functions \Eq{fragmentation_function}
    %
    (see also \Fig{shower:cartoon:fragmentation}).
    %
    We derived the relevant expressions for parton-to-parton fragmentation in \Sec{p2p-fragmentation}, where we found that the Mellin moments of parton-to-parton fragmentation functions were given by \Eq{ll_fragmentation_function}.
    %
    We repeat them here with \(\theta_f = 2\sqrt{\chi}\) (from \(\chi = (1-\cos\theta)/2, \theta\ll1\)) for convenience:
    \begin{align}
        \hat{F}_{f \leftarrow i} (m;\,
        \theta_f = 2 \sqrt{\chi}\leftarrow \Rjet)
        =
        \le(
            \alpha_s\le(2\,Q\sqrt{\chi}\ri)
            \, / \,
            \alpha_s\le(Q\,\Rjet\ri)
        \ri)
        ^
        {2 \hat{p}(m) \, / \, \beta_0}
        _{f\leftarrow i}
        \,;
    \end{align}

@@

Futhermore, since \Eq{jet_fn_ll_0} gives the (normalized) contribution of a narrow jet of partonic flavor \(i\) to the EEC in the collinear limit, it is precisely the jet function of \Eq{jet_fn_defn}.
%
Using \Eqss{ll_fragmentation_function}{splitting_fn_contribution_to_eec}{jet_fn_ll_0}, we can then write the LL jet function (EEC for a narrow jet of flavor \(i\)) as
\begin{align}
    \label{eq:eec_jet_fn_ll_1}
    J^i(\chi; \, Q)
    &=
    \frac{2 \as}{\chi}
    \,
    \sum_{a,b,\,f}
    \le(
        \int_0^1 \dd z \, z (1-z) \,
        P_{ab \leftarrow f}(z)
    \ri)
    \,
    \le(
        \frac{\alpha_s\le(2\,Q\sqrt{\chi}\ri)}
        {\alpha_s\le(Q\,\Rjet\ri)}
    \ri)
    ^
    {\frac{2\hat{p}(2)}{\beta_0}}
    _{f\leftarrow i}
    .
\end{align}

@@

Unfortunately, the expression of \Eq{old_def} is computationally expensive due to the intricate phase space constraint \(R_L = \max_{k,\ell}\{R_{i_k,i_\ell}\}\).
%
For example, the time required to compute the (integer) \gls{penc} scales as $M^N/N!$ for a jet with $M$ particles; when analytically continued to non-integer values of $N$, the \gls{penc} suffers  a computational scaling of $2^{2M}$.%
%
\footnote{
    \textsc{FastEEC}~\cite{Budhraja:2024xiq} achieves a substantial speed up by replacing particles with subjets whose radius is chosen dynamically, depending on the desired level of angular resolution.
    %
    For non-integer $N$, \textsc{FastEEC} furthermore uses a recursive algorithm that reduces the standard $2^{2M}$ scaling down to $M\,2^M$~\cite{Budhraja:2024tev}.
}
%
This computational cost impedes several exciting \gls{penc} applications.
%
For example, \glspl{penc} have an approximate scaling behavior related to the $N$-th Mellin moment of the DGLAP splitting functions~\cite{Chen:2020vvp};
%
the analytic continuation of the \gls{penc} to non-integer $N$ therefore provides access both to the full splitting functions of QCD and, in the limit $N \to 0$, a unique opportunity to study small-$x$ physics and BFKL dynamics with jets~\cite{Chen:2020vvp,Neill:2020bwv,Budhraja:2024tev}.
%
At very large values of \(N\), \glspl{penc} also encode additional fundamental features of QCD~\cite{Chen:2020vvp,Dai:2024wff}, such as level crossings with twist-4 operators.
%
Furthermore, in the high-multiplicity environment of heavy ion collisions, even computing \glspl{penc} for \(N=3\) has been computationally challenging, limiting their potential in studying medium effects~\cite{Bossi:2024qho}.
%
Improved computational efficiency is therefore necessary to leverage the full potential of \glspl{penc} in the study of realistic data samples.

Therefore, rather than focusing on the traditional projections of ENCs, in this section we focus on new parametrizations of higher point energy correlators, introduced in \Reff{Alipour-fard:2024szj}, with a number of improved properties.
%
First, our parametrization of the \gls{penc} depends on the largest distance \(R_1\) to a ``special'' particle $s$ in a set of $N$ particles, suitably averaged over all choices for $s$;
%
this yields simpler phase space restrictions than the traditional parametrization for the \gls{penc} in terms of the largest pairwise angle~\cite{Chen:2020vvp}.
%
Second, when considering more differential information, our parametrization of resolved ENCs (\gls{renc}s) employs non-redundant polar coordinates centered around the special particle, as in \Fig{cartoon}, yielding intuitive visualizations of jet substructure.
%
This differs from the traditional approach, which uses over-complete information from the set of all pairwise distances and neglects information about the relative orientation of particles.
%
Third, our parametrization offers dramatic improvements in computational performance, which is essential for their use in experimental analyses.
%
Finally, we anticipate that these conceptual and computational improvements will yield simpler theoretical calculations.
%
Our implementation for computing the \glspl{penc} of \Eq{new_def} be found on GitHub as an update to \textsc{FastEEC}~\cite{Budhraja:2024xiq}, and of our \glspl{penc} and \gls{renc}s at \texttt{ResolvedEnergyCorrelators}~\cite{github:RENC}.

@@

EWOCs therefore trade the ability to capture correlations between all particles in a jet for the ability to more easily control the effects of low-energy QCD.
%
As we will see, this tradeoff dramatically increases the richness of the correlations that can be probed by EWOCs.
As we discuss further in \Sec{ewoc-irc}, EWOCs are not collinear safe without an infra-red regulator;
%
therefore, we consider EWOCs which capture correlations between \textit{subjets}.%
\footnote{
    In principle, we could use any infra-red regulator -- see Remark~\ref{rem:lund-eec}.
}
%
These subjet EWOCs, which we will henceforth simply call EWOCs, have some important similarities and differences from the energy correlators defined above.

In addition to the angular correlations probed by the EEC, EWOCs can be used to capture many other correlations -- such as correlations in mass, formation time, or any other observable on subjets -- which are forbidden by collinear safety in the absence of a subjet radius.
%
An EWOC for a generic observable \(\mathcal{O}(\text{subjet 1, subjet 2})\) on pairs of subjets takes a similar form to the form of the EEC given in \Def{eec}.
%
Its definition was given in \Eq{intro_ewoc_def} and is repeated here for convenience:
%
\begin{definitionbox}{Pairwise EWOC}{pairwise-ewoc}
    A \glslink{ewoc}{\vocab{pairwise (or two-point) EWOC}} takes the form
    \begin{align}
        \label{eq:ewoc_def_2}
        \frac{\dd \Sigma_\mathcal{O}}{\dd \chi}
        =
        \frac{1}{\sigma}
        \int \dd\sigma \,
        \sum_{
            \text{subjets }
            i,\, j
        } \,
        z_i \, z_j \,
        \delta\left(\chi \, - \, \mathcal{O}_{ij}\right)
        ,
    \end{align}
    where \((\dd)\sigma\) indicates the (differential) cross section for jet production in the process under consideration and the sum on \(i\) and \(j\) run over the subjets within the jet.
    %
    The energy-weighting factors \(z_{i,j}\), adopted from the EEC, ensure that higher-energy subjets contribute more to the EWOC of \Eq{ewoc_def_2}.%
    \footnote{
        For proton-proton collisions at the LHC, transverse momentum fractions \(z_i = p_{T,i}/p_{T,\text{jet}}\) are a natural choice, while energy fractions \(z_{i} = E_{i}/E_\text{jet}\) are more natural in the context of electron-positron collisions.
    }
    %
    \(\mathcal{O}_{ij}\) is a user-defined  observable on pairs of subjets, such as their angular separation or invariant mass.
\end{definitionbox}

@@

\remark{}{
    For example, using the two-particle angular separation in \Eq{ewoc_def_2}, \(\mathcal O_{ij} \to \theta_{ij}\), recovers the traditional EEC of $e^+e^-$ collisions in the limit of zero subjet radius (i.e.~where the sum over pairs of subjets is replaced by a sum over pairs of particles).
}

@@

The advantage of EWOCs is that they directly probe correlations in an observable of interest.
%
The choice of subjet definition (e.g.~radius and recombination scheme) determines the set of subjets that contribute to the EWOC, and therefore may also be tuned to capture different physical effects.
%
For example, the subjet radius may be chosen to capture the subjets emerging from the hadronic decays of a boosted particle, as we explore in \Sec{ewoc-mass}.
%
In this section, we focus on the concrete example of the mass EWOC as a proof-of-concept, with a subjet radius chosen to probe decays of the \(W\) boson, but the use of EWOCs is by no means limited to the measurement of mass correlations.
%
Possible pairwise observables include:
\begin{itemize}
    \item
        The opening angle \(\theta_{ij}\) or the rapidity-azimuth distance \(R_{ij}\) between subjets, which probe angular correlations (used in $e^+e^-$ or $pp$ collisions, respectively).
        %
        These observables lead to the EEC, for which the use of subjets does not provide novel results;%
        \footnote{
            With a subjet radius \rsub{}, the value of the subjet EEC is nearly identical to the particle-level EEC for $\chi>\rsub$, while the collinear cutoff of the subjet radius sets the subjet EEC to zero for $0<\chi<\rsub$.
        }

    \item
        The mass \(m_{ij} = \sqrt{\le(p_i + p_j\ri)^2}\), which probes mass scales%
        \footnote{
            The four-momenta of the subjets depends on the recombination scheme used to define the subjets.
            %
            In this section, we use winner-take-all schemes which enforce that all subjets are massless.
            %
            We have found that this choice of recombination scheme does not affect the performance of the mass EWOC in constraining the $W$-boson mass.
        }, applied to the phenomenolgy of massive decaying particles in \Sec{ewoc-mass};
